Estimating the replicability of psychology experiments after an initial failure to replicate

### 2024-03-20 manuscript/manuscript.Rmd
finish editing pass

::: {.abstract data-latex="" lang="en"}
When a replication fails, scientists have to decide whether to make a second attempt or move on. Researchers who attempt to replicate studies before building on them often face this decision, given the relatively low rates of replication success in psychology. Here, we report on 17 re-replications of experiments with failed replications reported in Boyce et al. (2023). In 5/17 of these "rescue" projects (29%), the re-replication mostly or fully replicated the original results, albeit with smaller effect sizes. While a few replications were rescued with larger sample sizes and minor methodological changes, most re-replications failed again. In many situations, it may be most efficient to stop pursuing an effect after a single failed replication.
:::
TODO FIX CITATION / BIB INFO

@@

Imagine the scenario where you are a new graduate student, and you run a replication of a study in the literature that you are interested in building upon in your research.
The replication fails: perhaps the interaction you hoped for is directionally correct, but the point estimate is small and the confidence interval definitely overlaps 0, with a p-value of .3.
Or perhaps the interaction is numerically in the wrong direction, and also the main effects look different.
Alternatively the main effect of interest may be null, but you found a significant effect interaction that wasn't hypothesized.
Whatever the details of the replication failure, you are left with a question: Should you *try again* and run a re-replication, or should you *give up* and pick a different study to replicate and build on?

In psychology, large scale replication projects have tended to find replicability rates around half.
Across 100 studies with positive results, RP:P replicated 36 - 47% depending on the metric for replication success [@openscienceconsortium].
With large multi-site samples, Many Labs 1 replicated 11/13 effects (84.6%), Many Labs 2 replicated 14/28 effects (50%), and Many Labs 3 replicated 3/10 effects (30%) [@klein2014; @klein2018; @ebersole2016].
@camerer2018 included 21 behavioral social science studies and replicated 12-14 of them (57%-67%) depending on the metric used.
@boyce2023 reports an average replication score of 49% for 176 replications primary in psychology.
Psychology is not the only discipline where not everything replicates.
Large scale replication projects in other disciplines have found 39/97 (40%) for positive effects in cancer biology [@errington2021], 11/18 (61%) in economics [@camerer2016], and 31/40 (78%) for experimental philosophy [@cova2021].

Replications can fail for any reason. In principle, the question of why a replication failed is the same as asking why did two closely related studies have different results, which could be due to any differences between the study. In the case of replication failure, there's an asymmetry where the original study happened first and set the methods and analyses and got the more significant result.  [VB is struggling to articulate ideas here] 
One source of replication failure is that the result is fragile in some way: it could be a statistically unlikely result that was achieved through chance or p-hacking, or it could be sensitive to the exact conditions and time it was run under ("hidden moderators"). The reported effect sizes might be biased estimates due to some combination of heterogeneity, p-hacking, luck, underpowering, and publication bias. The replication study could be underpowered to detect the effect, perhaps due to powering for the reported (inflated) effect size, or unexpected attrition in the replication sample.

Replications might get different results from an original study due to differences in how the question of interest was instantiated; these differences in design could be intentional adaptations, or they could be unintentional if a replication recreates instructions or materials because the original ones are not available, or they could be trivial differences that turn out to be crucial to the effect.

A difference in experiment platforms (ex. online versus in-person) might mean that an implementation doesn't fit the platform, or the data quality controls and attention checks are inadequate. If there are differences in time, place, or subject population between the two studies, there may need to be differences in the materials, instructions, or procedure to provide a good *translation* -- too many changes could cause differing results, but so could not enough changes.

When replications fail, we generally do not know why they failed, although we often speculate on potential reasons. [TODO not sure how to transition here] The goal of most large-scale replication studies has been metascientific, with the aims of estimating the true proportion of effects in a literature or understanding the role of heterogeneity. For these purposes, it may not matter exactly why any individual replication failed because they are assessed as a whole.

But when individual scientists attempt to replicate an effect with the goal of building on it, they do care about the individual failed replication. In an uncertain literature, we might not want to commit resources to building on an effect without first checking that we, with our resources and participant sample, can get the core effect we wish to build on. Thus, it's very important whether the replication failure was caused by something fixable about the replication or whether it was caused by something not-fixable about the original result (ex. extreme lack of generalizability, p-hacking, etc).

While one can theorize after the fact about potential causes of replication failure, in most circumstances it is hard to diagnose for certain what made for the difference in the results. An exception might be if a specific error in one study or the other is found, such that the results were not what was thought (analytic error) or the data was not what was thought (experimental error). Thus, we usually can't know whether the replication failure was caused by factors that could be feasibly addressed or not.

Re-replication could serve as a way of triangulating on reasons for replication failure, as well as another chance for a scientist to get a working protocol that picks up the effect of interest. Thus, scientists may want to know both how often and when a re-replication is likely to work and detect an effect that is indicative that future experiments in this tradition are likely to be informative.

<!--TODO THIS PARAGRAPH NEEDS HELP Factors that suggest another failure include heterogeneity and lack-of-generalizability ("hidden moderator"), as the second replication is likely to be closer to the first replication than to the original in these random factors.
TODO CITE THAT PAPER about types of heterogeneity and which way they would fall An original result due to p-hacking or a lucky false positive is unlikely to re-replicate.
Sensitivity to environment or not working well online are unlikely to be fixable if the replications remain constrained to using online methods.
Similarly, if the result is specific to a special population, then a new replication is also likely to fail, unless it has gained access to this population.
Factors that may be fixable include sample size issues with the replication, assuming the re-replication can recruit a larger sample or reduce attrition.
It may be possible to fix underpoweredness and inflated effect size, at least to an extent.
If the effect size was inflated and the first replication powered for the effect, a better powered (but not huge) replication may be able to recover a smaller, but still convincing effect.
Failures due to differences between the materials may or may not be fixable.
If the original materials are still unavailable, a second replication is no better off.
If a second replication can get closer to the original, either by using closer materials, or better adapting the study to a new environment, the second replication might succeed at better approximating the results of the original.
A failed manipulation check failure may not be fixable, but it might at least be diagnoseable.-->

Re-replication has not been studied much, so it is uncertain how definitive a replication failure is. 
One study on re-replications was @bersole2020 which re-replicated 10 psychology studies that had failed replications reported in RP:P [@openscienceconsortium]. 
After  @openscienceconsortium published their results with a replication rate around 40%, @gilbert2016 raised concerns that methodological differences were the reason for the low replication rate. In 11 of the OSF studies, concerns about protocol fidelity had been raised by original authors prior to data-collection (and thus not contingent on results).
Of these 11 with concerns, 10 failed to obtain significant results in the RP:P replication [@openscienceconsortium]. In a follow-up, @ebersole2020 re-replicated 10 of these 11 in larger samples, using both the RP:P protocol and a new protocol revised under advice of the original authors or other experts, thus testing whether larger sample sizes and/or "better" methodologies would "rescue" the failed RP:P replications.

The result: 0 of the RP:P protocols found significant results, 2 of the 10 revised protocols did (but not the 1 that had a significant result in the original RP:P sample).
The replication effect size was much smaller CITE NUMBER [@ebersole2020].
We could frame @ebersole2020 as an attempt to recover the original effects of the 9 studies with significant original effects and non-significant RP:P effects by a combination of high power and design tweaks.
This suggests a successful re-replication rate of 2/9 (22%) for studies with non-significant first replications, under fairly favorable, high resource circumstances.

After a failed replication, we might speculate on what caused the replication result, and what that would mean for future work on the topic.
As the @gilbert2016 and @ebersole2020 example showed, these speculations on causes and thus re-replication outcomes are not always correct.

While @ebersole2020 looked at re-replication success rate in a high resource setting with access to large, multi-site samples and expert input, most early career scientists do not have those kinds of resources to devote to re-replication attempts.
On the other hand, early career researchers may also be prone to a different distribution of causes of initial replication failure.

Here, we investigate re-replication in the context of graduate student replications.
Students in a graduate methods class on experimental methods were learning about best practices for experimental research.
One option for their class project was to "rescue" an experiment that a student in a prior year had previously failed to replicate (reported in @boyce2023).
Thus, students were likely to be looking at both the original paper and the failed replication through the lens of best practices to see what they could fix.
[TODO Mike mentioned something about a sampling side box on experimentology that VB is having trouble finding]

In the present paper, we report the results of `r num_proj` re-replication "rescue" projects which each re-replicated a study which had a failed or only partially-successful replication in @boyce2023. 
TODO PREVIEW RESULTS??

@@

Our analysis plan was pre-registered after students had selected projects, but before final data collection on the projects.
Each project was also individually pre-registered by the student conducting it.
The overall analysis is at LINK, individual pre-registrations are linked from LINK.

We note one deviation from our pre-registration here: we pre-registered visual comparisons between original, first replication, and rescue projects using prediction intervals.
Prediction intervals depend on both the original effect size and variance CITATION TODO, and also the variance of the comparison (replication or rescue) study.
Thus we cannot show *the* prediction interval for the original study, but would have to show a prediction interval between each pair of studies, which we thought would not offer clarity.

@@

The experiments that were re-replicated were chosen from studies that failed to replicate in @boyce2023.
We created an initial list of `r nrow(filtered)` rescue-eligible studies that had received a subjective replication success score of 0, .25 or .5 (on a 0-1 scale) in @boyce2023, where the replication had a github repository available (github repositories were used starting in academic year 2015-2016), and where the original experiment had 200 or fewer participants (to ensure we could afford to increase the sample size).
We then contacted the replication project authors for permission to share their report and github repository with a new student and include it as a supplement on a resulting paper.
This left `r length(approved)` options for the students to choose from.
`r length(chosen)` students chose to do rescue projects.
`r length(chosen)-num_proj` students took an incomplete or did not indicate interest in being part of this paper, leaving a final sample of `r num_proj` rescue projects.

@@

Students were required to think of reasons the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions TODO LINK was given to students.
In general, we encouraged students to add manipulation checks as appropriate, and better adapt materials to online studies.
For instance, the rescue of @paxton2012 switched from using the CRT which is overused in online samples for the newer TODO.
The rescue of @jara-ettinger2022 discovered that the replication had accidentally used the drawn version of the stimuli rather than the photographic stimuli used in the original and reverted to the photographic stimuli.
@tarampi2016 original had participants indicate their answers (left or right) on a piece of paper in a timed navigation task, the replication had them indicate by clicking a drop down, and rescue went with pressing keyboard keys to indicate the direction of the a turn.
TODO others to note?
Once students experimental designs and analytic plans were approved by TAs (VB and BP), students pre-registered and ran their samples.

@@

Each project was rated on the basis of subjective replication success both by MCF and by one of VB and BP.
Disagreements were resolved through discussion.
As a compliment to the subjective rating of overall success, we stastically compared one key measure of interest for each study, following @boyce2023. In order to statistically compare the key measures, we needed effect sizes reported in the same way for each original study, first replication, and rescue.
When effects were not reported in consistent ways across original and replications, we recalculated effects from raw data when necessary to obtain comparable values.

We also recorded the same set of potential correlates that were used in @boyce2023 for original, first replication, and rescue (these were already rated for original and first replication).
These potential correlates included the subfield of the study (cognitive, social, or other psychology), its publication year, experimental design features including whether it was a within- or between- subject design, whether each condition was instantiated with one vignette or multiple, how many items each participant saw, and whether there were open materials and open data.

For the original study and each replication, we recorded the number of participants post-exclusions.
For studies where some extra conditions were dropped, we count only the participants in the key conditions all replications had for comparability.
For instance, if an original study compared between two critical conditions but also had a baseline control, we would not count the participants in the baseline condition if a replication did not include this condition.
We also recorded whether each study was conducted online with a crowdsourced platform or not.

@@

Our primary question of interest is how many of these `r num_proj` rescue projects succeeded at replicating the results in the original study.
When a replication fails to obtain the same results, one may have intuitions about what may have gone wrong -- these rescue projects test how often addressing these potential issues in fact works.

@@

In the original replication sample from @boyce2023, `r boyce2023_sample |> filter(sub_rep>.5) |> nrow()` out of `r boyce2023_sample |> nrow()` replications (`r round(percent_rep*100)`%) mostly or fully replicated (i.e. received a subjective replication score of .75 or 1).
Note that @boyce2023 report the average replication score as a percent success (49%), but given that we considered studies with a subjective score of .5 as eligible to be rescued, we recomputed the success rate when scores of 0-.5 as failures and .75 and 1 as successes.
If the re-replication rate in our sample is representative of the re-replication rate for the initially non-replicating studies, then the combined chance of mostly or fully replicating in a first replication or one follow-up replication is `r round(overall*100)`%.

@@

[TODO Ben: I think the paragraph below could use more explanation about what $p$-original is, but I'm not confident enough in my understanding of it to write it myself.]

Here, we measure statistical consistency using $p$-original, the $p$-value on the null hypothesis that two effects come from the same distribution.
As our primary comparison, we compare the effect size of the original study to the meta-analytic effect of the totality of the replications (first replication, rescue, and additional if found).
Because there is only a small number of replications per study, we impute the heterogeneity value of $\tau=.21$ SMD, which is the average level of heterogeneity found by @olsson-collentine2020 in prior multi-site replications in psychology.

[ query: should we bootstrap any of the following?
]

The median value of $p$-original between an original study and its replications was `r main_summ$median |> round(2)` [IQR: `r main_summ$low |> round(2)` - `r main_summ$high |> round(2)`].
`r round(main_summ$less_than*100)`% of the $p$-original values were less than .05, indicating by conventional thresholds a rejection of the null hypothesis that the original and the replications came from the same distribution for the given imputed level of heterogeneity.
Individual $p$-original values for each study are shown in Table \@ref(tab:tab-porig).

As secondary measures, we calculated the $p$-original values between a) the original and the rescue, b) the original and non-rescue replications, and c) the rescue and other replications.
For a) the rescue versus the original, median value of $p$-original was `r rescue_summ$median |> round(2)` [IQR: `r rescue_summ$low |> round(2)` - `r rescue_summ$high|> round(2)`], and `r round(rescue_summ$less_than*100)`% of the $p$-original values were less than .05.
For b) the non-rescue replications versus the original, median value of $p$-original was `r rep_summ$median |> round(2)` [IQR: `r rep_summ$low |> round(2)` - `r rep_summ$high |> round(2)`], and `r round(rep_summ$less_than*100)`% of the $p$-original values were less than .05.
For c) the rescue versus the other replications, median value of $p$-original was `r rescue_rep_summ$median |> round(2)` [IQR: `r rescue_rep_summ$low |> round(2)` - `r rescue_rep_summ$high |> round(2)`], and `r round(rescue_rep_summ$less_than*100)`% of the $p$-original values were less than .05.

@@

Overall, allowing for a heterogeneity level of $\tau$=.21 SMD, a number of rescues and replications are not statistically consistent with the original effects.
The pattern of inconsistency does not align with which studies were rated as having replicated.
In all cases, the point estimate of the re-replication is smaller than, or in the opposite direction of, the original effect on the key measure of interest.

@@

We ran correlations between the set of predictor variables used in @boyce2023 and the subjective replication scores of the rescues.
We also added some (non-preregistered) predictors related to the sample size of the first replication, after seeing the successful re-replications of @ngo2019 and @krauss2003, both studies with small replication samples.

All of the correlations are presented in Figure \@ref(tab:cor-table).
As the number of rescues is small, and many of these predictors are correlated, we caution against over-interpretation.
The strongest correlates of rescue success were open materials, a small sample size on the first replication, a small sample size on the first replication relative to the original sample size, and a large rescue sample size relative to the first replication.
None of the pre-registered correlates meet the conventional significance threshold; the two correlates based on ratios that reflect a relatively smaller replication sample size are marginally significant.

@@

An additional factor that influences the interpretation of a replication is how close the replication's methods were to the original.
In the rescue projects, we aimed to have methods be as close as was feasible or appropriate.
(As discussed above, changes in time, population, or setting may necessitate adaptation of a study to maximize fidelity.) However, rescue projects varied in how close the re-replications actually were, often due to limitations in the availability of original stimuli and original instructions, in addition to the use of primarily online subject pools.
Table \@ref(tab:tab-size) shows the closeness of each first replication and rescue according to the classification scheme from @lebel2018. TODO CLOSENESS NUMBERS SHOULD BE DOUBLE CHECKED AND FIRST REPS REVISED AS NEEDED

@@

# Case studies

@@

Another successful rescue was that of @ngo2019.
Here, the original study found a large effect.
The first replication, powering for 80% power on the reported effect, recruited a small sample of 12 people.
It failed to find the effect.
The rescue, powered using 2.5x the original sample (as recommended by @simonsohn2015), recovered a clear effect (albeit a much smaller one).
There are reasons to think that some effect sizes in the literature may be inflated TODO CITATIONS, and separate reasons that slight changes to experiments or switches to online could result in noisier samples and thus smaller effect sizes.
Therefore, replications with smaller samples than the original (even if powered to the original effect size), may not be very diagnostic, and could potentially benefit from a re-replication.

@@

(could discuss tarampi as an example where there were potential issues, we fixed them and it still didn't work?)

@@

The rescues all showed smaller effect sizes than their original studies, regardless of whether the pattern of effects replicated.
A large minority of replications had effect sizes that were statistically inconsistent with the original effect, even allowing for some heterogeneity between studies.
These diminished and inconsistent effects suggest that even if a re-replication "works", it may be difficult to build upon as follow-up studies will need large samples to detect small effects.

@@

The authors of the rescue projects put substantial effort into trying to set up rescues that had a good chance of success, but projects were constrained by budget limitations, a short timeline, and primarily running online studies.
These limitations are representative of the sort of resource limitations often faced by early-career researchers.
That said, it is possible that different results might be obtained in better-resourced settings, by scientists with more expertise and more time.
Thus, we do not make statements about whether the original results are "true" or "false-positives", we merely claim they do not support cumulative research by early-career researchers under the constrained conditions we tested.

We opened this paper with a question about what an early-career psychology researcher should do given a failed replication: should they try again or move on?
From our sample of testing the "try again" approach, it seems that the odds of a re-replication working are low (consistent with @ebersole2020).
Especially if there is not a clear, identifiable reason for the first replication's failure, another try is unlikely to recover the original result.

@@

>>>>

::: {.abstract data-latex="" lang="en"}
When a replication fails, scientists have to decide whether to make a second attempt or move on. Psychology researchers who attempt to replicate studies before building on them often face this decision, given the relatively low rates of replication success in psychology. Here, we report on 17 re-replications of experiments with failed replications reported in Boyce et al. (2023). In 5/17 of these "rescue" projects (29%), the re-replication mostly or fully replicated the original results, albeit with smaller effect sizes. While a few replications were rescued with larger sample sizes and minor methodological changes, most re-replications failed again. In many situations, it may be most efficient to stop pursuing an effect after a single failed replication.
:::
TODO FIX CITATION / BIB INFO (re OSC for RP:P but only after we unlink from zotero)

TODO compile everything into OSF repo & link
TODO link individual pre-reg's in repo
TODO link overall pre-reg
TODO include in repo & link the reasons for replication failure that were offered to students
TODO redo/double check closeness scores for first rep and rescues 
TODO add something about Verity's to case-study? 
TODO author contributions, acknowledgements

@@

Imagine the scenario where you are a new graduate student, and you run a replication of a study in the literature that you are interested in building upon in your research.
The replication fails: perhaps the interaction you hoped for is directionally correct, but the point estimate is small and the confidence interval definitely overlaps 0, with a p-value of .3.
Or perhaps the interaction is numerically in the wrong direction, and also the main effects look different.
<!--Or maybe the main effect of interest may be null, but you found a significant effect interaction that wasn't hypothesized.-->
Whatever the details of the replication failure, you are left with a question: Should you *try again* and run a re-replication, or should you *give up* and pick a different study to replicate and build on?

In psychology, large scale replication projects have found that around half the studies successfully replicate. 
Across 100 studies with positive results, RP:P replicated 36 - 47% depending on the metric for replication success [@openscienceconsortium].
With large multi-site samples, Many Labs 1 replicated 11/13 effects (84.6%), Many Labs 2 replicated 14/28 effects (50%), and Many Labs 3 replicated 3/10 effects (30%) [@klein2014; @klein2018; @ebersole2016].
@camerer2018 included 21 behavioral social science studies and replicated 12-14 of them (57%-67%) depending on the metric used.
@boyce2023 reported an average replication score of 49% for 176 replications primary in psychology.
Psychology is not the only discipline where many studies do not replicate.
Large scale replication projects in other disciplines have found 39/97 of studies with positive effects (40%) replicating in cancer biology [@errington2021], 11/18 studies (61%) replicating in economics [@camerer2016], and 31/40 studies (78%) replicating in experimental philosophy [@cova2021]. Across these disciplines, and especially in psychology, large-scale replications indicate a substantial chance of scientists encountering failed replications if they try to replicate published findings.

Replications can fail for many reasons. In principle, any difference between the original study and replication could be a root cause of the difference in their outcomes. In practice, the relationship is not symmetric, and many causes of replication failure can be attributed more to the original study or more to the replication. 
One potential reason for replication failure is that the original result is fragile in some way: it could be a statistically unlikely result that was achieved through chance or p-hacking, or it could be sensitive to the exact conditions and time it was run under ("hidden moderators"). Another possibility is that the reported effect size might be inflated due to some combination of heterogeneity, p-hacking, low power, and publication bias. One consequence of inflated effect sizes is that replication studies might power for the reported effect size and then have too small a sample to reliably detect the (true) smaller effect. Replication studies might also end up underpowered due to unexpected attrition or noise in the replication sample.

Replications could also get different results from an original study due to differences in how the question of interest was instantiated. These differences in design could be intentional adaptations, unavoidable changes from lack of access to the original instructions or materials, or changes that were not expected to be critical. A difference in experiment platforms (ex. online versus in-person) might mean that an implementation no longer is appropriate, or the data quality controls and attention checks are inadequate. If there are differences in time, place, or subject population between the two studies, corresponding changes to the materials, instructions, or procedure may be needed to provide a good *translation* -- too many changes could cause differing results, but so could not enough changes.

When replications fail, we generally do not know why they failed, although we may speculate. For the metascientific goals of most large-scale replication projects, it doesn't matter why any particular study failed to replicate, because the aim is to estimate the true proportion of effects in a literature.

In contrast, when individual scientists attempt to replicate an effect with the goal of building on it, they do care about why an individual replication failed. In an uncertain literature, scientists might not want to commit resources to building on an effect without first checking that they, with their resources and participant sample, can get the core effect. In this case, it's very important whether the replication failure is due to something fixable about the replication or something not-fixable about the original result.

While one can theorize after the fact about potential causes of replication failure, in most circumstances it is hard to diagnose for certain what made for the difference in the results. An exception might be if a specific error in one study or the other is found, such that the results were not what was thought (analytic error) or the data was not what was thought (experimental error). Aside from these rare cases, we usually can't know whether the replication failure is fixable just by looking at the replication failure itself.

Instead, re-replication could serve as a way of triangulating on reasons for re-replication failure, as well as another chance for a scientist to get a working protocol that picks up the effect of interest. Thus, scientists may want to know what the chances are that a re-replication succeeds at detecting an effect that indicates future experiments with this protocol are worth pursuing.

Re-replication has not been studied much, so it is uncertain how often re-replications recover the results of an original study. That is, it is not well known how definitive a first replication failure is.

One estimate of re-replication success rates comes from @ebersole2020 which re-replicated 10 psychology studies that had failed replications reported in RP:P [@openscienceconsortium]. 
After @openscienceconsortium published their results with a replication rate around 40%, @gilbert2016 raised concerns that methodological differences were the reason for the low replication rate. In 11 of the RP:P studies, the original authors had raised concerns about protocol fidelity prior to data-collection.
Of these 11 with concerns, 10 failed to obtain significant results in the RP:P replication [@openscienceconsortium]. In a follow-up, @ebersole2020 re-replicated 10 of these 11 in larger samples, using both the RP:P protocol and a new protocol revised under advice of the original authors or other experts, thus testing whether larger sample sizes and/or "better" methodologies would "rescue" the failed RP:P replications.

The result was that 0 of the RP:P protocols found significant results, and 2 of the 10 revised protocols did (but not the 1 that had a significant result in the original RP:P sample). Even when a significant effect was recovered, the effect size was much diminished [@ebersole2020].
We could frame @ebersole2020 as an attempt to recover the original effects of the 9 studies with significant original effects and non-significant RP:P effects by a combination of high power and design tweaks.
This suggests a successful re-replication rate of 2/9 (22%) for studies with non-significant first replications, under fairly favorable, high resource circumstances.

While it is tempting to speculate on the reasons for failed replications, the @gilbert2016 and @ebersole2020 example shows these speculations on causes of replication failure do not always pan out in re-replication outcomes.

While @ebersole2020 looked at re-replication success rate in a high resource setting with access to large, multi-site samples and expert input, most early career scientists do not have those kinds of resources to devote to re-replication attempts. On the other hand, early career researchers may also be prone to a different distribution of causes of initial replication failure. Thus, it is an open question what the re-replication rate is for failed replications of early career researchers, and estimating this rate could inform decisions on how to act after failed replications.

Here, we investigate re-replication in the context of graduate student replication class projects.
Students in a graduate methods class on experimental methods were learning about best practices for experimental research and had the option to "rescue" an experiment that a student in a prior year had previously failed to replicate as their class project (first replications reported in @boyce2023).
Thus, students looked at both the original paper and the failed replication through the lens of best practices to see what potential issues in the replication they could fix. 
In the present paper, we report the results of `r num_proj` re-replication "rescue" projects which each re-replicated a study which had a failed or only partially-successful replication in @boyce2023. We find that a minority of projects (5/17, 29%) successful recovered the original effect.

@@

Our analysis plan was pre-registered after students had selected projects, but before final data collection on the projects.
Each project was also individually pre-registered by the student conducting it.
The overall analysis is at LINK, individual pre-registrations are linked from LINK.
We note one deviation from our pre-registration here: we pre-registered visual comparisons between original, first replication, and rescue projects using prediction intervals.
Prediction intervals depend on both the original effect size and variance and the variance of the comparison (replication or rescue) study [@patil2016].
Thus we cannot show *the* prediction interval for the original study, but would have to show a prediction interval between each pair of studies, which we thought would not offer clarity.

@@

The experiments that were re-replicated were chosen from studies that failed to replicate in @boyce2023.
We created an initial list of `r nrow(filtered)` rescue-eligible studies that had received a subjective replication success score of 0, .25 or .5 (on a 0-1 scale) in @boyce2023, where the replication had a Github repository available (github repositories were used starting in academic year 2015-2016), and where the original experiment had 200 or fewer participants (to ensure we could afford to increase the sample size).
We then contacted the replication project authors for permission to share their report and repository with a new student and include it as a supplement on a resulting paper.
This left `r length(approved)` options for the students to choose from.
`r length(chosen)` students chose to do rescue projects, and
`r length(chosen)-num_proj` students took an incomplete or did not indicate interest in being part of this paper, leaving a final sample of `r num_proj` rescue projects.

@@

Students were required to think of reasons the first replication might not have worked, and address them if they could.
A list of possible reasons and solutions TODO LINK was given to students.
In general, we encouraged students to add manipulation checks as appropriate, and better adapt materials to online studies.
For instance, the rescue of @paxton2012 switched from using the CRT which many online participants are overfamiliar with for the newer CRT-2 which is less overused but gets at the same construct [@thomson2016].
The rescue of @jara-ettinger2022 discovered that the replication had accidentally used the drawn version of the stimuli rather than the photographic stimuli used in the original and reverted to the photographic stimuli.
@tarampi2016 had participants do a timed navigation task where they wrote L or R to indicate which way to turn at each intersection on a paper map. The replication ported this experiment online and had participants click a drop down to select left or write for each turn. The rescue switched for the unwieldy dropdown menus to having participants press keys on the keyboard to indicate the direction of a turn, which seemed like a more natural interface.
Once students experimental designs and analytic plans were approved by TAs (VB and BP), students pre-registered and ran their samples.

@@

Each project was rated on the basis of subjective replication success both by MCF and by one of VB and BP.
Disagreements were resolved through discussion.
As a complement to the subjective rating of overall success, we statistically compared one key measure of interest for each study, following @boyce2023. In order to statistically compare the key measures, we needed effect sizes reported in the same way for each original study, first replication, and rescue.
When effects were not reported in consistent ways across original and replications, we recalculated effects from raw data when necessary to obtain comparable values.

We recorded the same set of potential correlates that were used in @boyce2023 for original, first replication, and rescue (these were already rated for original and first replication).
These potential correlates included the subfield of the study (cognitive, social, or other psychology), its publication year, experimental design features including whether it was a within- or between- subject design, whether each condition was instantiated with one vignette or multiple, how many items each participant saw, and whether there were open materials and open data.

For the original study and each replication, we recorded the number of participants post-exclusions.
For studies where some extra conditions were dropped in some replications, we counted only the participants in the key conditions all replications had for comparability.
For instance, if an original study compared between two critical conditions but also had a baseline control, we did not count the participants in the baseline condition if a replication did not include this condition.
We also recorded whether each study was conducted online using crowd-sourced platform or not.

@@

Our primary question of interest was how many of the `r num_proj` rescue projects succeeded at replicating the results in the original study.
When a replication fails to obtain the same results, one may have intuitions about what may have gone wrong -- these rescue projects test how often addressing these potential issues in fact works.

@@

In the original replication sample from @boyce2023, `r boyce2023_sample |> filter(sub_rep>.5) |> nrow()` out of `r boyce2023_sample |> nrow()` replications (`r round(percent_rep*100)`%) mostly or fully replicated (i.e. received a subjective replication score of .75 or 1).
Note that @boyce2023 report the average replication score as a percent success (49%), but given that we considered studies with a subjective score of .5 as eligible to be rescued, we recomputed the success rate where scores of 0, .25, and .5 are considered failures and .75 and 1 successes.
If the re-replication rate in our sample is representative of the re-replication rate for the initially non-replicating studies, then the combined chance of mostly or fully replicating in a first replication or one follow-up replication is `r round(overall*100)`%.

@@

Here, we measure statistical consistency using $p$-original, the $p$-value on the null hypothesis that one study's population effect comes from the same distribution of population effects as another study or studies [@mathurNewStatisticalMetrics2020a] If we assume no heterogeneity between studies, then this is the same as asking how consistent one study's estimate of the population-level effect is with another study's estimate of the population-level effect. However, meta-analyses of closely related studies, such as those from different sites of multi-site replications, indicate that there is often heterogeneity between studies. We cannot accurately estimate the level of heterogeneity for an effect meta-analytically as we have only a small number of replications per study. Instead, we impute a heterogeneity value of $\tau=.21$ SMD, which is the average level of heterogeneity found by @olsson-collentine2020 in prior multi-site replications in psychology.

As our primary comparison, we compare the effect size of the original study to the meta-analytic effect of the totality of the replications (first replication, rescue, and additional if found). The value of $p$-original thus represents how likely an effect size equal to or more extreme than the original effect size is to occur if it comes from a distribution of population effect sizes defined by the mean and standard error of the replications (combined meta-analytically) with a between-population standard deviation of $\tau=.21$. That is, it $p$-original is a measure of the statistical consistency of the original result with the totality of the replications, assuming a given level of heterogeneity of effects between studies. Smaller $p$-original values indicate more inconsistency.

The median value of $p$-original between an original study and its replications was `r main_summ$median |> round(2)` [IQR: `r main_summ$low |> round(2)` - `r main_summ$high |> round(2)`].
`r round(main_summ$less_than*100)`% of the $p$-original values were less than .05, indicating by conventional thresholds a rejection of the null hypothesis that the original comes from the same distribution of population effects as the replications. 
Individual $p$-original values for each study are shown in Table \@ref(tab:tab-porig).

As secondary measures, we calculated the $p$-original values between a) the original and the rescue, b) the original and non-rescue replications, and c) the rescue and other replications. 
For a) the original versus the rescue, median value of $p$-original was `r rescue_summ$median |> round(2)` [IQR: `r rescue_summ$low |> round(2)` - `r rescue_summ$high|> round(2)`], and `r round(rescue_summ$less_than*100)`% of the $p$-original values were less than .05.
For b) the original versus all the non-rescue replications, median value of $p$-original was `r rep_summ$median |> round(2)` [IQR: `r rep_summ$low |> round(2)` - `r rep_summ$high |> round(2)`], and `r round(rep_summ$less_than*100)`% of the $p$-original values were less than .05.
For c) the rescue versus the other replications, median value of $p$-original was `r rescue_rep_summ$median |> round(2)` [IQR: `r rescue_rep_summ$low |> round(2)` - `r rescue_rep_summ$high |> round(2)`], and `r round(rescue_rep_summ$less_than*100)`% of the $p$-original values were less than .05.

@@

Overall, allowing for a heterogeneity level of $\tau$=.21 SMD, a number of original effects are not statistically consistent with rescues and replication effects. 
The pattern of inconsistency does not align with which studies were rated as having replicated.
In all cases, the point estimate of the re-replication is smaller than, or in the opposite direction of, the original effect on the key measure of interest.

@@

Are there signals in the features of a study or it's replication that predict whether a re-replication will succeed? To address this question, we ran correlations between the set of predictor variables used in @boyce2023 and the subjective replication scores of the rescues.
We also added some (non-preregistered) predictors related to the sample size of the first replication, after seeing the successful re-replications of @ngo2019 and @krauss2003, both of which had small replication samples.

The number of rescues is small, and many of these predictors are correlated, so we caution against over interpretation. All of the correlations are presented in Figure \@ref(tab:cor-table).
The strongest correlates of rescue success were open materials, a small sample size on the first replication, a small sample size on the first replication relative to the original sample size, and a large rescue sample size relative to the first replication.
None of the pre-registered correlates meet the conventional significance threshold; the two correlates based on ratios that reflect a relatively smaller replication sample size are marginally significant.

@@

An additional factor that influences the interpretation of a replication is how close the replication's methods were to the original.
In the rescue projects, we aimed to have methods be as close as was feasible or appropriate. However, rescue projects varied in how close the re-replications actually were, often due to limitations in the availability of original stimuli and original instructions, in addition to the use of primarily online subject pools.
Table \@ref(tab:tab-size) shows the closeness of each first replication and rescue according to the classification scheme from @lebel2018. TODO CLOSENESS NUMBERS SHOULD BE DOUBLE CHECKED AND FIRST REPS REVISED AS NEEDED

@@

## Case studies

@@

Another successful rescue was that of @ngo2019.
Here, the original study found a large effect.
The first replication, powering for 80% power on the reported effect, recruited a small sample of 12 people.
It failed to find the effect.
The rescue, powered using 2.5x the original sample (as recommended by @simonsohn2015), recovered a clear effect (albeit a much smaller one).
There are reasons to think that some effect sizes in the literature may be inflated [@openscienceconsortium, @camerer2018], and it is also possible that changes to experiments or switches to online could result in noisier samples and thus smaller effect sizes.
Therefore, replications with smaller samples than the original (even if powered to the original effect size), may not be very diagnostic, and could potentially benefit from a re-replication.

@@

@@

The rescues all showed smaller effect sizes than their original studies, regardless of whether the pattern of effects replicated.
A large minority of replications had effect sizes that were statistically inconsistent with the original effect, even accounting for expected levels of heterogeneity between studies.
These diminished and inconsistent effects suggest that even if a re-replication "works", it may be difficult to build upon as follow-up studies will need large samples to detect small effects.

@@

The authors of the rescue projects put substantial effort into trying to set up rescues that had a good chance of success, but projects were constrained by budget limitations, a short timeline, and primarily running online studies.
These limitations are representative of the sort of resource limitations often faced by early-career researchers.
That said, it is possible that different results might be obtained in better-resourced settings, by scientists with more expertise and more time.
Thus, where re-replications failed, we do not make any statements about whether the original results were false positives, we merely claim the original results do not support cumulative research by early-career researchers under the constrained conditions we tested.

We opened this paper with a question about what an early-career psychology researcher should do given a failed replication: should they *try again* or *give up* and move on?
From our sample of testing the *try again* approach, it seems that the odds of a re-replication working are low (consistent with @ebersole2020).
Especially if there is not a clear, identifiable reason for the first replication's failure, another try is unlikely to recover the original result, and it may be more efficient to *give up* and build on a different result instead.

@@

TODO
